A Formal Model of Checked C

### 2023-01-15 sandp23/appendix.tex
finished appx.

\subsection{Compilation Formalism}\label{sec:compilation}

@@

The main subtlety of compiling \checkedc to Clang/LLVM is to capture the annotations on pointer literals
that track array bound information, which is used in premises
of rules like \textsc{S-DefArray} and
  \textsc{S-AssignArr} to prevent spatial safety violations.
The \checkedc compiler \cite{li22checkedc} inserted additional pointer checks 
for verifying pointers are not null and the bounds are within their limits. \lc{what does it mean by bounds within limits?}
The latter is done by introducing additional shadow variables for storing (NT-)array pointer bound information.

@@

In \systemname, context and pointer modes determine the particular heap/function store that a pointer points to,
i.e., $\cmode$ pointers point to checked regions, while $\tmode$ and $\umode$ pointers point to unchecked regions. \lc{context mode doesn't determine where pointer points to but check if it is valid to point to that region?}
Unchecked regions are associated with a sandbox mechanism that permits exception handling of potential memory failures.
In the compiled LLVM code, pointer access operations have different syntaxes when the modes are different. \lc{should we explain why we use different syntax instead of compiling the corresponding heap to llvm}
\Cref{fig:flagtable} lists the different compiled syntaxes of a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang). The columns represent different pointer modes and the rows represent context modes.
For example, when we have a $\tmode$-mode pointer in a $\cmode$-mode region, we compile a deference operation to the sandbox pointer access function ($\texttt{sand\_get}(x)$) accessing the data in the \systemname implementation. In \lang, we create a new deference data-structure on top of the existing $\estar{x}$ operation (in LLVM): $\getstar{m}{x}$. If the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.

This section shows how \lang deals with pointer modes, mode switching and function pointer compilations, 
with no loss of expressiveness
as the \checkedc contains the erase of annotations in \cite{li22checkedc} and \Cref{appx:comp1}.
For the compiler formalism, 
we present a compilation algorithm that converts from
\lang to \elang, an untyped language without metadata
annotations, which represents an intermediate layer we build on LLVM for simplifying compilation. 
In \elang, the syntax for deference, assignment, malloc, function calls are: $\getstar{m}{e}$, $\elassign{m}{e}{e}$, 
$\emalloc{m}{\omega}$, and $\elcall{m}{e}{\overline{e}}$. \lc{strlen?}
The algorithm sheds
  light on how compilation can be implemented in the real Checked C
  compiler, while eschewing many vital details (\elang has many 
  differences with LLVM IR).

%This section shows how \systemname deals with 
%annotations can be safely erased: using static information a compiler
%can insert code to manage and check bounds metadata, with no loss of
%expressiveness. We present a compilation algorithm that converts from
%\lang to \elang, an untyped language without metadata
%annotations. The syntax and semantics \elang
  %closely mirrors that of \lang; it differs only in that literals lack
  %type annotations and its operational rules perform no
  %bounds and null checks, which are instead inserted during
  %compilation. Our compilation algorithm is evidence that \lang's
  %semantics, despite its apparent use of fat pointers, faithfully
  %represents Checked C's intended behavior. The algorithm also sheds
  %light on how compilation can be implemented in the real Checked C
  %compiler, while eschewing many important details (\elang has many 
  %differences with LLVM IR).

Compilation is defined by extending \lang's
typing judgment as follows:
\[\Gamma;\Theta;\rho \vdash_m e \gg \dot e:\tau\]
There is now a \elang output $\dot e$ and an input $\rho$, which maps
each (NT-)array pointer variable to its mode and
each variable \code{p} to a pair of \emph{shadow
  variables} that keep \code{p}'s up-to-date upper and lower bounds. 
These may differ from the bounds in \code{p}'s type due to bounds
widening.\footnote{Since lower bounds are never widened, the
  lower-bound shadow variable is unnecessary; we include it for uniformity.} 
\lc{These two sentences are confusing. }

@@

\subsection{Approach}

Due to space constraints, we explain the rules for compilation by
examples, using a C-like syntax; the complete rules are given in
\iftr
Appendix~\ref{appx:comp1}.
\else
the supplemental report~\cite{checkedc-tech-report}.
\fi
Each rule performs up to three tasks: (a) conversion of $e$ to
A-normal form; (b) insertion of dynamic checks and bound widening expressions; 
and (c) generate right pointer accessing expressions based on modes.
%
A-normal form conversion is straightforward: compound expressions are managed by storing results of subexpressions into temporary variables,
as in the following example.

@@

\myparagraph{Pointer Accesses and Modes}
%
In every declaration \lc{(or the beginning of a function body) why?} of a pointer,
if the poniter is an (NT-)array,
we first allocate two \emph{shadow variables}
to track the lower and upper bounds which are potentially changed for pointer arithmetic and NT-array bound widening.
Each $\cmode$-mode NT-array pointer variable is associated with its type information in a store.
Additionally, we place bounds and null-pointer checks, such as the line 6 and 7 in \Cref{fig:compilationexample1}.
In addition, in the formalism, before every use of a tainted pointer (\Cref{fig:compilationexample1} line 9 and 10), 
there is an inserted verification step similar to \Cref{fig:const-type},
which checks if a pointer is well defined in the heap (\code{not_null}) and the spatial safety.
Predicate \code{not_null} checks that every element in the pointer's range (\code{p_lo} and \code{p_hi}) is well defined in the heap.  
The modes in compiled deference (\code{*(mode(p) }$\wedge$\code{ m,p)})
 and assignment (\code{*(mode(q) }$\wedge$\code{ m,q)=1}) operations 
are computed based on the meet 
operation ($\wedge$) of the pointer mode (e.g. \code{mode(p)}) and the current context mode (\code{m}).

\myparagraph{Checked and Unchecked Blocks}
%
In the \systemname implementation,
$\euncheckedtext$ and $\echeckedtext$ blocks 
are compiled as context switching functions provided by the sandbox mechanism.
$\eunchecked{\overline{x}}{e}$ is compiled to 
$\texttt{sandbox\_call}(\overline{x},e)$, where we call the sandbox 
to execute expression $e$ with the arguments $\overline{x}$.
$\echecked{\overline{x}}{e}$ is compiled to 
$\texttt{callback}(\overline{x},e)$, where we perform 
a \texttt{callback} to a checked block code $e$ inside a sandbox.
In \systemname, we adopt an aggressive execution scheme that
directly learns pointer addresses from compiled assembly to make the $\texttt{callback}$ happen.
In the formalism, we rely on the type system to 
guarantee the context switching without creating the extra function calls for simplicity.

@@

\myparagraph{Function Pointers and Calls}
%
Function pointers are managed similarly to normal pointers,
but we insert checks to check if the pointer address is not null in 
the function store instead of heap, and whether or not the type is correctly represented, 
for both $\cmode$ and $\tmode$ mode pointers 
\footnote{$\cmode$-mode pointers are checked once in the beginning and $\tmode$-mode pointers are checked every time when use}.
For example, in compiling the \code{stringsort} function in \Cref{fig:checkedc-example-1},
we place a check \code{verify_fun(cmp, not_null(c, p_lo, p_hi) && type_match)},
and we place a similar check before \Cref{fig:checkedc-example-3} line 7 to check the tainted \code{cmp} when it is used. 
The compilation of function calls (compiling to $\elcall{m}{e}{\overline{e}}$) 
is similar to the manipulation of pointer access operations in \Cref{fig:flagtable}.

For compiling dependent function calls,
\Cref{fig:compilationexample1} provides a hint.
Notice that the bounds for the array pointer \code{p} are not passed as
arguments. Instead, they are initialized according to \code{p}'s
type---see line~4 of the original \lang program at the top of the figure.
Line~$3$ of the generated code
sets the lower bound  to \code{0} and the
upper bound to \code{n}.

\subsection{Metatheory}
\label{sec:meta}

% \review{ 
% My interpretation of section IV.C is that the authors have a pen-and-paper
%   proof of Theorem 4, and that random testing using the Redex models has been
%   used to gain confidence in such a proof. Is this correct?}
% \liyi{Yes, this is correct.}
% \mwh{No, not correct: We have no pen-and-paper proof, just the
%   model. Clarify here}
% showing that
% extensive random testing fails to falsify the bisimilarity
% property. \mwh{Better way to state the previous?}

@@

\ignore{
 We formulate our simulation theorem differently
because the standard simulation theorem imposes a very strong
syntactic restriction to the compilation strategy. Very often, $(\dot
\varphi_0, \dot \heap_0, \dot e_0)$ reduces to a term that is
semantically equivalent to $(\dot \varphi_1, \dot \heap_1, \dot e_1)$,
but we are unable to syntactically equate the two configurations due
to the extra binders generated for dynamic checks and ANF
transformation. In earlier versions of the Redex model, we attempted
to change the compilation rules so the configurations could match
syntactically. However, the approach scaled poorly as we added
additional rules. 
This slight relaxation on the equivalence relation
between target configurations allows us to specify compilation more
naturally without having to worry about syntactic constraints.
}

@@

The current version of the Redex model has been tested against $23000$
expressions with depth less than $11$. Each expression can
reduce multiple steps, and we test simulation between every two
adjacent steps to cover a wider range of programs, particularly the
ones that have a non-empty heap.

    \path [->] (A) edge node {$\longrightarrow_\cmode$} (B);
    \path [<->] (A) edge node {$\gg$} (C);
    \path [<->] (B) edge node {$\gg$} (D);
    \path [dashed,<->] (C) edge node {$\sim$} (D);
    \path [dashed,->] (C) edge node {$\xrightarrow{\cdot}^*$} (E);
    \path [dashed,->] (D) edge node[above] {$\xrightarrow{\cdot}^*$} (E);
\end{scope}

% \ignore
% {The two theorems can be best understood by the diagram in Fig.~\ref{fig:checkedc-simulation-ref}.
% % The diagram also indicates the simulation relation ($\sim$) among our formalization of the \checkedc Type System ($\vdash$), Semantics ($\longrightarrow$), Compilation ($\gg$), as well as the compiled C semantics ($\xrightarrow{c}$).
% On the top of the relation, we have type soundness theorem (Sec.~\ref{sec:theorem}) stating that every type-checked \lang program $e_0$ can transition via \lang semantics into a type-checked program $e_1$ by taking a checked step.
% The second line refers to that both $e_0$ and $e_1$ can be compiled to a C program $\dot e_0$ and $\dot e_1$, and such compilations always exist. The bottom two rewrites ($\xrightarrow{c}$) state that both $\dot e_0$ and $\dot e_1$ are transitioned finally to the same state $\dot e$.

%  The simulation states that given any well-typed \lang expression $e_0$, its compiled C program as $\dot e_0$, if $e_0$ can transition one step to $e_1$, and there must exist the compiled C program $\dot e_1$, such that both $\dot e_0$ and $\dot e_1$ transition to the same final state $\dot e$ in C. The simulation relation $\sim$ is built on top of the "triangle" structure in Fig.~\ref{fig:checkedc-simulation-ref} stating that every \lang expression and its one step transitioned expression might be compiled to two different C program, but the C semantics always evaluates them to the same place.
% The translation preservation theorem is a support for the simulation theorem stating that for any well-typed \lang expression $e_0$,  its one step \checkedc transition expression $e_1$ must exist a compiled C program through the compilation process ($\gg$).

% A corollary of the simulation theorem states that for any \lang expression $e_0$ and its one step \lang transition expression $e_1$, such that $e_0 \longrightarrow e_1$, if $e_0$ contains a fault due to a unchecked block (the blame theorem tells us that if there is a fault in a \lang code, it must come from a unchecked block), and its compiled code $\dot e_0$ is evaluated to a fault state, the compiled code $\dot e_1$ for $e_1$ is always evaluated to the same fault state in C.
% This is an important property for a \lang compiler to preserve because any problematic program will be captured by running it long enough, so that enough random testing cases are able to capture all the bugs.
% In our formalization implementation, the type soundness theorem is proved through Coq, while the simulation theorem is validated through intensive random test case generation. We generate tens of thousands cases and cover all possible corner cases of the compilation. 
% }
% \ignore{
% \begin{figure*}[t]
%   \begin{subfigure}[b]{1\textwidth}
%     \begin{lstlisting}
% int deref_array (size_t n, nt_array_ptr<int> p : bounds(p, p + n)) {
%   return *p;
% }

% ...
% /* nt_array_ptr<int> p0 : bounds(p0, p0 + 5) */
% deref_array(5, p0);
%     \end{lstlisting}
%     \label{fig:chkcexamplederef}
%     \caption{\texttt{deref\_array} in \checkedc}
%   \end{subfigure}

%   \begin{subfigure}[b]{1\textwidth}
% \begin{lstlisting}
% int deref_array(size_t n, int *p) {
%   /* statically determine the definitions p_lo and p_hi */
%   int *p_lo = p;
%   int *p_hi = p + n;
%   /* runtime checks */
%   assert(p_lo <= p && p <= p_hi);
%   assert(p != NULL);
%   /* widening */
%   int p_derefed = *p;
%   if (p_derefed != '\0') {
%     if (p_hi == p) {
%       ++p_hi;
%     }
%   }
%   return p_derefed;
% }
% ...
% /* int *p0 */
% /* rho(p0) = p_lo, p_hi */
% deref_array(5, p0);
% \end{lstlisting}
% \caption{\texttt{deref\_array} in C}
%     \label{fig:cexamplederef}
%   \end{subfigure}

%   \begin{subfigure}[b]{\textwidth}
%     \begin{lstlisting}
% /* nt_array_ptr<int> p : bounds(e0, e1) */
% /* rho(p) = p_lo, p_hi */
% size_t x = strlen(p);
% if(x >= 1) {
%   *(p+1);
% }
%     \end{lstlisting}
%     \caption{\texttt{strlen} in \checkedc}
%     \label{fig:chkcexamplestrlen}
%   \end{subfigure}

%   \begin{subfigure}[b]{\textwidth}
%     \begin{lstlisting}
% /* nt_array_ptr<int> p : bounds(e0, e1) */
% /* rho(p) = p_lo, p_hi */
% /* runtime checks omitted */
% ...
% size_t x = strlen(p);
% int *p_hi_new = p + x;
% p_hi = max(p_hi, p_hi_new);
% if (x >= 1) {
%   /* null check for pointer arithmetic */
%   assert(p != NULL);
%   int *p_1 = p + 1;
%   /* null check for dereference */
%   assert(p_1 != NULL);
%   /* bounds check for dereference */
%   /* note how we are using p_hi rather than p + x */
%   assert(p_lo <= p_1 && p_1 <= p_hi);
%   ...
% }
%     \end{lstlisting}
%     \caption{\texttt{strlen} in C}
%     \label{fig:cexamplestrlen}
%   \end{subfigure}
% \caption{Compilation of null-terminated string dereference}
% \label{fig:compilationexample}
% \end{figure*}
% }

% \ignore{
% Fig.~\ref{fig:compilationexample} gives an example of how we compile the
% dereference of null-terminated strings by inserting explicit checks and
% bounds widening code. For readability, we present the example in real
% \checkedc and C syntax. At
% line 3-4 of Fig.~\ref{fig:cexamplederef}, we see how the upper and lower
% bounds are defined in terms of the arguments \code{p} and \code{n}
% according to the bounds annotations obtained during typechecking. This
% avoids the need to pass in bounds as extra arguments, thus maintaining
% compatibility with C code. When we call \code{deref_array} on
% \code{p0}, a string with size 5 (excluding the null character), there
% is no need to pass the bounds \code{p_lo} and \code{p_hi} stored on
% the stack. It is possible that \code{p_hi} has been widened
% before we reach line 21. The programmer will have to perform a
% \code{dynamic_bounds_cast} to recover the more precise bounds information. The if statement from line 10 to line 14
% attempts to widen the bound when the dereferenced result is not
% null. \todo[inline]{Show a full-fledged C program?} The widened upper bound will be available within the scope of
% \code{deref_array}, in contrast to the T-IfNT rule, which only
% remembers the widened bounds within the scope of the then branch of
% the if statement.

% \todo[inline]{Widening can happen at every
%   dereference at runtime. Is that ok?} Fig.~\ref{fig:cexamplestrlen}
% shows how an invocation of \code{strlen} on null-termianted strings
% is compiled into C code. We perform the same runtime checks that
% happen during dereferencing. The widening code at line
% 8 updates \code{p}'s upper bound only if the result of
% \code{strlen} is larger than the value of the upper bound stored on
% the stack. This is another scenario where the runtime can be more precise
% than the statically determined bounds information.

% }

\subsection{Additional Type and Semantic Rules}\label{appx:add-type-sem}

Since bounds expressions may
contain variables, determining assumptions like $b_l \leq b_l'$
requires reasoning about the probable values of these variables'. The type
system uses $\Theta$ to make such reasoning more precise.
$\Theta$ is a map from variables $x$ to
equation predicates $P$, which have the form $P ::= \tgez \;|\; \teq{b}$.
It maps variables to equations that are recorded along the type checking procedure.
If $\Theta$ maps $x$ to $\tgez$, that means that $x \ge 0$;
$\teq{b}$ means that $x$ is equivalent to the bound value $b$ in the current context, 
such as in the type judgment for $e_2$ in Rule \textsc{T-LetInt} and \textsc{T-RetInt}.
\iftr
  Appendix~\ref{app:le}.
\else
  the supplemental report~\cite{checkedc-tech-report}.
\fi has an example rule for populating $\Theta$ with a $\tgez$ predicate.

\begin{figure}[t]
{\small
{\captionsetup[lstlisting]{margin = 8 mm}
  \begin{lstlisting}[xleftmargin=8 mm]
nt_array_ptr<char> safe_strcat
   (nt_array_ptr<char> dst : count(n),
    nt_array_ptr<char> src : count(0), int n) {
  int x = strlen(dst);
  int y = strlen(src);
  nt_array_ptr<char> c : count(n) =
    dyn_bounds_cast
           <nt_array_ptr<char>>(dst,count(n));
    // sets c == dst with bound n (not x)
  if (x+y < n) {
    for (int i = 0; i < y; ++i)
      *(c+x+i) = *(src+i);
    *(c+x+y) = '\0';
    return dst;
  }
  return null;
}
  \end{lstlisting}
}
}
\caption{Implementation of safe \code{strcat}}
\label{fig:strcat-ex}
\end{figure}

Recall that array
bounds in types may refer to in-scope variables; e.g., parameter
\code{dst}'s bound \code{count(n)} refers to parameter \code{n} on lines
2-3 in \Cref{fig:strcat-ex}. 
All dependent bound variables appearing in function headers 
are bounded by a variable list in the function pointer type, which is checked in the type rule.
Semantically,
the call is expanded into a \texttt{let} which binds
parameter variables $\overline{x}$ to the actual arguments
$\overline{n}$, but annotated with the parameter types
$\overline{\tau}$ (this will be safe for type-correct programs). 
The function body $e$ is wrapped in a static cast
$(\tau[\overline{n} / \overline{x}])$ which is the function's return
type but with any integer parameter variables $\overline{x}$ appearing in that
type, as type bound variables,
substituted with the call's actual arguments $\overline{n}$.
To see why this is needed, suppose that \code{safe_strcat} in
Fig.~\ref{fig:strcat-ex} is defined to return a
\code{nt_array_ptr<int>:count(n)} typed term, and assume that we
perform a \code{safe_strcat} function call as
\code{x=safe_strcat(a,b,10)}. After the evaluation of \code{safe_strcat}, the
function returns a value with type \code{nt_array_ptr<int>:count(10)}
because we substitute bound variable \code{n} in the 
defined return type with \code{10} from the function call's
argument list.

For compiling dependent function calls,
\Cref{fig:compilationexample1} provides a hint.
Notice that the bounds for the array pointer \code{p} are not passed as
arguments. Instead, they are initialized according to \code{p}'s
type---see line~4 of the original \lang program at the top of the figure.
Line~$3$ of the generated code
sets the lower bound  to \code{0} and the
upper bound to \code{n}.

\begin{defi}[Type Environment Well-formedness]\label{type-wellformed}
A type environment $\Gamma$ is well-formed if every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
\end{defi}

\begin{defi}[Heap Well-formedness]
For every $m$, A heap $\heap$ is well-formed if (i) $\heap(m,0)$ is undefined, and
(ii) for all $\evalue{n}{\tau}$ in the range of $\heap(m)$, type $\tau$
contains no free variables. 
\end{defi}

\begin{defi}[Stack Well-formedness]
A stack snapshot $\varphi$ is well-formed if
for all $\evalue{n}{\tau}$ in the range of $\varphi$, type $\tau$
contains no free variables. 
\end{defi}

\begin{defi}[Stack Consistency]
A type environment $\Gamma$, variable predicate set $\Theta$, and
stack snapshot $\varphi$ are consistent---written $\Gamma;\Theta\vdash
\varphi$---if for every variable $x$, $\Theta(x)$ is defined implies
$\Gamma(x) = \tau$ for some $\tau$ and 
$\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq_{\Theta} \tau$. 
\end{defi}

\begin{defi}[Checked Stack-Heap Consistency]
A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---if
for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ with $\mode(\tau)=\cmode$ implies $\emptyset;\heap(\cmode);\emptyset \vdash_{\cmode} n:\tau$.
\end{defi}

\begin{defi}[Checked Heap-Heap Consistency]
A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---if
for every constant $n$, $\emptyset;\heap;\emptyset \vdash_{\cmode} n:\tau$ implies $\emptyset;\heap';\emptyset \vdash_{\cmode} n:\tau$.
\end{defi}

\myparagraph{Constant Validity}
Rules \textsc{T-ConstU} and \textsc{T-ConstC} in \Cref{fig:type-system-1}
describe type assumptions for constants appearing in a program.
$\neg \cmode(\tau)$ judges that a constant pointer 
in an unchecked region cannot be of a checked type.
The restriction ensures that programmers 
cannot guess a checked pointer address and utilize it in an unchecked region in \systemname.
In rule \textsc{T-ConstC}, we requires a static 
verification procedure for validating a constant pointer in \Cref{fig:const-type}.

The verification process $\Theta;\heap;\sigma \vdash_m n : \tau$
validates the constant $\evalue{n}{\tau}$, 
where $\heap(m)$ is the initial heap that the constant resides on and
$\sigma$ is a set of constant assumed to be checked.
A global function store $\Xi(m)$ is also required to check the validity of a function pointer.
A valid function pointer should appear in the right store region ($\cmode$ or $\umode$)
and the address stores a function with the right type.
The last rule in \Cref{fig:const-type} describes the validity check for a non-function pointer, 
where every element in the pointer range ($[0,\size(\omega))$) should be well
typed.
A checked pointer checks validity in type step as rule \textsc{T-ConstC},
while a tainted/unchecked pointer does not check for such during the type checking.
Tainted pointers are validated through the validity check in dynamic execution as we mentioned in rule \rulelab{S-DefT}.

\subsection{Meta Theories}\label{sec:theorem}

% Before we present our main theorems, we need to first
% discuss the meaning what a pointer being well-typed in a given heap
% snapshot $\heap$ means, which is captured by rules in
% Fig.~\ref{fig:const-type}. The variable type rule ($\textsc{T-Var}$)
% simply checks if a given variable has the defined type in $\Gamma$;
% the constant rule ($\textsc{T-Const}$) is slightly more involved.
% First, it ensures that the type annotation $\tau$ does not contain any
% free variables. More importantly, it ensures that the pointer points
% to a location that makes sense in a given heap.
%  
%  
%  The $\size$ function in Fig.~\ref{fig:const-type}
% refers to the \code{sizeof} function in C computing the number of
% bytes for a type.
%  
%  
%  Second, we
% require that any constant ($\evalue{n}{\tau}$) should make sense in
% $\heap$. We develop a recursive predicate $\sigma \vdash n : \tau$ to
% verify if $n$ has $\tau$ in a heap snapshot $\heap$. $\sigma$ is a
% constant set containing the constants that have been verified by the
% relation. For every constant $\evalue{n}{\tau}$, it is either an
% integer $\tint$, an unchecked pointer $\tptr{\omega}{\umode}$,
% zero-valued number ($n=0$), checked in $\sigma$
% ($\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma$); or if it is not the
% above case, then (i) $\heap(n)$ is defined, and (ii) for every heap
% location $n+i$ in the range of the pointer (if $\omega$ is a word
% type, range is $[0,1)$; if $\omega$ is an array type
%   ($\tarray{0}{b_h}{\tau'}$), range is $[0,b_h)$, if $\tau$ is a
%     NT-array type ($\tntarray{0}{b_h}{\tau'}$), range is $[0,b_h+1)$),
%       if $\heap(n+i)=\evalue{n_a}{\tau_a}$, then
%       $\evalue{n_a}{\tau_a}$ satisfies $\sigma \cup \{(n,\tau) \}
%       \vdash n_a : \tau_a$.
%  
%  
% \begin{figure}[t]
% {\small
% \text{Type Rules for Constants and Variables:}
% \begin{mathpar}
%   \inferrule[T-Var]
%       {x : \tau \in \Gamma}
%       {\Gamma;\Theta \vdash_m x : \tau}
%  
%   \inferrule[T-Const]
%       {\fv(\tau) = \emptyset \\ \emptyset \vdash n : \tau}
%       {\Gamma;\Theta\vdash_m \evalue{n}{\tau} : \tau}
% \end{mathpar}
%     
% \text{Rules for Checking Constant Pointers In Heap:}
% \begin{mathpar}
%   \inferrule
%       {}
%       {\sigma \vdash n : \tint}
%  
%   \inferrule
%       {}
%       {\sigma \vdash n : \tptr{\omega}{\umode}}
%  
%   \inferrule
%       {}
%       {\sigma \vdash 0 : \tptr{\omega}{\cmode}}
%  
%   \inferrule
%       {\evalue{n}{\tptr{\omega}{\cmode}}\in \sigma}
%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
%  
%   \inferrule
%       {\forall i \in [0,\size(\omega)) .
%            \sigma \cup \{(n:\tptr{\omega}{\cmode}) \}\vdash \heap(n+i)}
%       {\sigma \vdash n : \tptr{\omega}{\cmode}}
% \end{mathpar}
% }
% \caption{Type Rules for Checking Constants/Variables}
% \label{fig:const-type}
% \end{figure}

% \review{
%  Theorem 1 refers to a program $e$ being well-formed. Unless I've missed
%   something, I didn't see such a definition in the paper.}
% \mwh{This was stale text (dropped); $e$'s well formedness follows from the
%   assumption of well typing; we have added more details about that.}

In this subsection, we focus on our main meta-theoretic results about
\lang: type soundness (progress and preservation),
non-exposure, and non-crashing.
These proofs have been conducted in our Coq model.
Type soundness relies on several \emph{well-formedness} given in \ref{li22checkedc} and \Cref{appx:add-type-sem}.
Progress states that a \lang program can always make a move:

\begin{thm}[Progress]\label{thm:progress}

>>>>

\subsection{Compilation Formalism}\label{appx:comp1}

@@

The main subtlety of compiling \checkedc to Clang/LLVM is to capture the annotations on pointer literals
that track array bound information, which is used in premises
of rules like \textsc{S-DefArrayC} and
  \textsc{S-AssignArrC} to prevent spatial safety violations.
The \checkedc compiler \cite{li22checkedc} inserted additional pointer checks 
for verifying pointers are not null and the bounds are within their limits.
The latter is done by introducing additional shadow variables for storing (NT-)array pointer bound information.

@@

In \systemname, context and pointer modes determine the particular heap/function store that a pointer points to,
i.e., $\cmode$ pointers point to checked regions, while $\tmode$ and $\umode$ pointers point to unchecked regions. 
Unchecked regions are associated with a sandbox mechanism that permits exception handling of potential memory failures.
In the compiled LLVM code, pointer access operations have different syntaxes when the modes are different.
\Cref{fig:flagtable} lists the different compiled syntaxes of a deference operation ($\estar{x}$) for the compiler implementation (\textsc{CBox}, stands for \systemname) and formalism (\textsc{Core}, stands for \lang). The columns represent different pointer modes and the rows represent context modes.
For example, when we have a $\tmode$-mode pointer in a $\cmode$-mode region, we compile a deference operation to the sandbox pointer access function ($\texttt{sand\_get}(x)$) accessing the data in the \systemname implementation. In \lang, we create a new deference data-structure on top of the existing $\estar{x}$ operation (in LLVM): $\getstar{m}{x}$. If the mode is $\cmode$, it accesses the checked heap/function store; otherwise, it accesses the unchecked one.

This section shows how \lang deals with pointer modes, mode switching and function pointer compilations, 
with no loss of expressiveness
as the \checkedc contains the erase of annotations in \cite{li22checkedc}.
For the compiler formalism, 
we present a compilation algorithm that converts from
\lang to \elang, an untyped language without metadata
annotations, which represents an intermediate layer we build on LLVM for simplifying compilation. 
In \elang, the syntax for deference, assignment, malloc, function calls are: $\getstar{m}{e}$, $\elassign{m}{e}{e}$, 
$\emalloc{m}{\omega}$, and $\ecall{e}{\overline{e}}$.
The algorithm sheds light on how compilation can be implemented in the real \systemname
  compiler, while eschewing many vital details (\elang has many differences with LLVM IR).

Compilation is defined by extending \lang's
typing judgment as follows:
\[\Gamma;\Theta;\rho \vdash_m e \gg \dot e:\tau\]
There is now a \elang output $\dot e$ and an input $\rho$, which maps
each (NT-)array pointer variable to its mode and
each variable \code{p} to a pair of \emph{shadow
  variables} that keep \code{p}'s up-to-date upper and lower bounds. 
These may differ from the bounds in \code{p}'s type due to bounds
widening.\footnote{Since lower bounds are never widened, the
  lower-bound shadow variable is unnecessary; we include it for uniformity.}

@@

\myparagraph{Approach}
%
Here, we explain the rules for compilation by
examples, using a C-like syntax; the complete rules are given in
\cite{checkedc-tech-report}.
Each rule performs up to three tasks: (a) conversion of $e$ to
A-normal form; (b) insertion of dynamic checks and bound widening expressions; 
and (c) generate right pointer accessing expressions based on modes.
%
A-normal form conversion is straightforward: compound expressions are managed by storing results of subexpressions into temporary variables,
as in the following example.

@@

\myparagraph{Pointer Accesses and Modes}
%
In every declaration of a pointer,
if the poniter is an (NT-)array,
we first allocate two \emph{shadow variables}
to track the lower and upper bounds which are potentially changed for pointer arithmetic and NT-array bound widening.
Each $\cmode$-mode NT-array pointer variable is associated with its type information in a store.
Additionally, we place bounds and null-pointer checks, such as the line 6 and 7 in \Cref{fig:compilationexample1}.
In addition, in the formalism, before every use of a tainted pointer (\Cref{fig:compilationexample1} line 9 and 10), 
there is an inserted verification step similar to \Cref{fig:const-type},
which checks if a pointer is well defined in the heap (\code{not_null}) and the spatial safety.
Predicate \code{not_null} checks that every element in the pointer's range (\code{p_lo} and \code{p_hi}) is well defined in the heap.  
The modes in compiled deference (\code{*(mode(p) }$\wedge$\code{ m,p)})
 and assignment (\code{*(mode(q) }$\wedge$\code{ m,q)=1}) operations 
are computed based on the meet 
operation ($\wedge$) of the pointer mode (e.g. \code{mode(p)}) and the current context mode (\code{m}).

\myparagraph{Checked and Unchecked Blocks}
%
In the \systemname implementation,
$\euncheckedtext$ and $\echeckedtext$ blocks 
are compiled as context switching functions provided by sandbox.
$\eunchecked{\overline{x}}{e}$ is compiled to 
$\texttt{sandbox\_call}(\overline{x},e)$, where we call the sandbox 
to execute expression $e$ with the arguments $\overline{x}$.
$\echecked{\overline{x}}{e}$ is compiled to 
$\texttt{callback}(\overline{x},e)$, where we perform 
a \texttt{callback} to a checked block code $e$ inside a sandbox.
In \systemname, we adopt an aggressive execution scheme that
directly learns pointer addresses from compiled assembly to make the $\code{_Callback}$ happen.
In the formalism, we rely on the type system to 
guarantee the context switching without creating the extra function calls for simplicity.

@@

\myparagraph{Function Pointers and Calls}
%
Function pointers are managed similarly to normal pointers,
but we insert checks to check if the pointer address is not null in 
the function store instead of heap, and whether or not the type is correctly represented, 
for both $\cmode$ and $\tmode$ mode pointers 
\footnote{$\cmode$-mode pointers are checked once in the beginning and $\tmode$-mode pointers are checked every time when use}.
The compilation of function calls (compiling to $\elcall{m}{e}{\overline{e}}$) 
is similar to the manipulation of pointer access operations in \Cref{fig:flagtable}.
For compiling dependent function calls,
\Cref{fig:compilationexample1} provides a hint.
Notice that the bounds for the array pointer \code{p} are not passed as
arguments. Instead, they are initialized according to \code{p}'s
type---see line~4 of the original \lang program at the top of the figure.
Line~$3$ of the generated code
sets the lower bound  to \code{0} and the
upper bound to \code{n}.

\subsection{Constraints and Metatheory}
\label{sec:meta}

Here, we first show some Well-formedness and consistency definitions that are required in \Cref{sec:theorem}, and then show the simulation theorem for the \lang compiler.
Type soundness relies on several \emph{well-formedness}:

\begin{definition}[Type Environment Well-formedness]\label{type-wellformed}
A type environment $\Gamma$ is well-formed if every variable mentioned as type bounds in $\Gamma$ are bounded by $\tint$ typed variables in $\Gamma$.
\end{definition}

\begin{definition}[Heap Well-formedness]
For every $m$, A heap $\heap$ is well-formed if (i) $\heap(m,0)$ is undefined, and
(ii) for all $\evalue{n}{\tau}$ in the range of $\heap(m)$, type $\tau$
contains no free variables. 
\end{definition}

\begin{definition}[Stack Well-formedness]
A stack snapshot $\varphi$ is well-formed if
for all $\evalue{n}{\tau}$ in the range of $\varphi$, type $\tau$
contains no free variables. 
\end{definition}

\begin{definition}[Stack Consistency]
A type environment $\Gamma$, variable predicate set $\Theta$, and
stack snapshot $\varphi$ are consistent---written $\Gamma;\Theta\vdash
\varphi$---if for every variable $x$, $\Theta(x)$ is defined implies
$\Gamma(x) = \tau$ for some $\tau$ and 
$\varphi(x) =\evalue{n}{\tau'}$ for some $n,\tau'$ where $\tau' \sqsubseteq_{\Theta} \tau$. 
\end{definition}

\begin{definition}[Checked Stack-Heap Consistency]
A stack snapshot $\varphi$ is consistent with heap $\heap$---written $\heap \vdash \varphi$---if
for every variable $x$, $\varphi(x)= \evalue{n}{\tau}$ with $\mode(\tau)=\cmode$ implies $\emptyset;\heap(\cmode);\emptyset \vdash_{\cmode} n:\tau$.
\end{definition}

\begin{definition}[Checked Heap-Heap Consistency]
A heap $\heap'$ is consistent with $\heap$---written $\heap \triangleright \heap'$---if
for every constant $n$, $\emptyset;\heap;\emptyset \vdash_{\cmode} n:\tau$ implies $\emptyset;\heap';\emptyset \vdash_{\cmode} n:\tau$.
\end{definition}

@@

@@

The current version of the Redex model has been tested against $21500$
expressions with depth less than $12$. Each expression can
reduce multiple steps, and we test simulation between every two
adjacent steps to cover a wider range of programs, particularly the
ones that have a non-empty heap.
